### 2011-04-22 standing-patches-paper.tex
latex bib oops
latex oops
bib works now?
minor notes

Suppose we fix our attention on a particular new mutation that happens to be the first to occur in some region.
As it spreads, it will cover a distance $L$ in time $L/v$, where $v = \sigma \sqrt{2s_b}$ is the speed of the wave of advance.
The number of other mutations appearing in the circle it has covered up until this time is Poisson with mean
$\lambda_0 \pi L^2 + \lambda \pi L^3 /v$ in two dimensions
(and $\lambda_0 2 L + \lambda \pi L^2 /v$ in one dimension).
% $\lambda_0 \omega_d L^d + \lambda \omega_d L^{d+1} /v$ in $d$ dimensions
Therefore, if we define $\chi$ to be the unique positive solution to
\[
    $\lambda_0 \pi \chi^2 + \lambda \pi \chi^3 /v = 1$,
\]
then $\chi$ gives the distance spread unobstructed by the descendants of a new mutant
before it is expected that one other successful mutation would have arisen in the area covered so far.
The explicit formula for $\chi$ is cumbersome but straightforward; here we omit the lengthy expression.
(This is in two dimensions; in one dimension the characteristic length is $\chi = \frac{ \sqrt{ 4 \lambda_0^2 + 2\lambda v } - 2 \lambda_0 }{ 4 \lambda v }$.)
This can be rewritten 
\begin{equation} \label{eqn:defines_chi}
    s_d^{-1} (\sqrt{2s_b} \chi)^2 + \sigma (\sqrt{2s_b} \chi)^2 = (\pi \rho \mu)^{-1} .
\end{equation}
From this we see that $\chi$ increases linearly with $\sqrt{s_b}$,
and that the relative sizes of $\sigma$ and $1/s_d$ determine the relative contributions of new and standing variants, respectively.
If $\sigma s_d \gg 1$, then standing variation is relatively unimportant, 
and this reduces to the characteristic length of \citet{ralphcoop2010}, $\chi_+ = (v/\lambda \pi)^{1/3}$.
On the other hand, if $\sigma s_d \ll 1$, then most adaptation come from standing variation, 
and the solution to \eqref{eqn:defines_chi} is close to $\chi_0 = (\lambda_0 \pi)^{-2}$. 
% More generally, as $\sigma s_d$ goes from $0$ to $\infty$, the positive solution to \eqref{eqn:defines_chi} moves continuously between the two.

@@

Describe simple discrete model.

Approximation in continuous model by branching Brownian motion with killing to find migration rate between patches -- 
good approximation neglecting interaction (if local population size is large). 
Look at Barton.

@@

Pull out results from other paper on discrete model.

when are patches effectively isolated?

>>>>

Suppose we fix our attention on a particular new mutation that happens to be the first to occur in some region.
As it spreads, it will cover a distance $L$ in time $L/v$, where $v = \sigma \sqrt{2s_b}$ is the speed of the wave of advance.
The number of other mutations appearing in the circle it has covered up until this time is Poisson with mean
$\lambda_0 \pi L^2 + \lambda \pi L^3 /v$ in two dimensions
(and $\lambda_0 2 L + \lambda \pi L^2 /v$ in one dimension).
% $\lambda_0 \omega_d L^d + \lambda \omega_d L^{d+1} /v$ in $d$ dimensions
Therefore, if we define $\chi$ to be the unique positive solution to
\[
    \lambda_0 \pi \chi^2 + \lambda \pi \chi^3 /v = 1,
\]
then $\chi$ gives the distance spread unobstructed by the descendants of a new mutant
before it is expected that one other successful mutation would have arisen in the area covered so far.
The explicit formula for $\chi$ is cumbersome but straightforward; here we omit the lengthy expression.
(This is in two dimensions; in one dimension the characteristic length is $\chi = \frac{ \sqrt{ 4 \lambda_0^2 + 2\lambda v } - 2 \lambda_0 }{ 4 \lambda v }$.)
This can be rewritten 
\begin{equation} \label{eqn:defines_chi}
    s_d^{-1} (\sqrt{2s_b} \chi)^2 + \sigma (\sqrt{2s_b} \chi)^2 = (\pi \rho \mu)^{-1} .
\end{equation}
From this we see that $\chi$ increases linearly with $\sqrt{s_b}$,
and that the relative sizes of $\sigma$ and $1/s_d$ determine the relative contributions of new and standing variants, respectively.
If $\sigma s_d \gg 1$, then standing variation is relatively unimportant, 
and this reduces to the characteristic length of \citet{ralphcoop2010}, $\chi_+ = (v/\lambda \pi)^{1/3}$.
On the other hand, if $\sigma s_d \ll 1$, then most adaptation come from standing variation, 
and the solution to \eqref{eqn:defines_chi} is close to $\chi_0 = (\lambda_0 \pi)^{-2}$. 
% More generally, as $\sigma s_d$ goes from $0$ to $\infty$, the positive solution to \eqref{eqn:defines_chi} moves continuously between the two.

@@

Describe simple discrete model, for general migration rates.

Get migration and mutation rates from a homogeneous continuous model:
The migration rate from patch $i$ to a patch at distance $r$ is $C_i \exp(-\gamma r \sqrt{s} /\sigma)$,
where $C_i$ depends on the size of the patch and (to a lesser degree?) the local structural details of the selective difference.
The mutation rate is more or less (size of patch) times $\mu$; 
more accurately we can integrate using Barton assuming step selection.

@@

Pull out results from other paper on discrete model:
if migration rates are equal then we get Chinese Restaurant/Ewens;
in any case is independent of selection.

when are patches effectively isolated?   
This occurs when the effective migration rate is much smaller than the mutation rate, namely...
